\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,geometry}
\geometry{margin=1in}
\title{The One-Equation Law for Reality-Aligned Agents}
\author{Andraž Đurič \and GPT-5 Thinking}
\date{}

\begin{document}
\maketitle
\vspace{-1.2em}

\noindent\textbf{Core law (single equation).}
\begin{equation}
\boxed{
\min_{P \in \mathcal{F}} \; \mathrm{CT}(P;P^\star)
\quad\text{with}\quad
\mathrm{CT}(P;P^\star)=\lambda\,D_{\mathrm{KL}}\!\big(P(W,M,S,G)\,\Vert\,P^\star(W,M,S,G)\big)
}
\end{equation}
\noindent\(W\)=world, \(M\)=models, \(S\)=signals/actions, \(G\)=objectives. Canonical scale \(\lambda=k_B T\ln 2\).
Ethical feasibility: \(\mathcal{F}=\{P:\mathbb{E}_P[H_{\text{severe}}]\le \varepsilon\}\).

\medskip\noindent\textbf{Design rule.}
Treat any controllable increase in CT as forbidden inefficiency, except when required to satisfy the ethical constraint.

\medskip\noindent\textbf{Invariants.}
Non-negativity (\(\mathrm{CT}\ge0\), equality iff \(P=P^\star\)); additivity over independent subsystems; monotone decrease under truthful refinement toward \(P^\star\); uniqueness (KL is the canonical functional under standard axioms).

\medskip\noindent\textbf{Operational loop.}
Maintain a versioned reference \(P^\star\) from evidence; drive the induced \(P\) toward \(P^\star\) within \(\mathcal{F}\); publish diffs and CT-proxy metrics (log-loss gaps, calibration, constraint violations).

\medskip\noindent\textbf{Falsifiable implications.}
(L) Corrections of inconsistency cost \(\ge \lambda\,\Delta D_{\mathrm{KL}}\).
(T) Transparency/andon-like stops reduce variance and tail risk at fixed resources.

\medskip\noindent\textbf{Scope.}
Applies to humans, software agents, and institutions. Minimal requirement: auditable \(P^\star\), measurable CT proxies, explicit ethics gate.

\end{document}
